---
title: "PAs vs PES in Brazil"
author: "Julia Lunardi"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Protected Areas vs Payment for Ecosystem Services in Brazil

This is the code I am developing for my Master's thesis about the effectiveness
of Protected Areas versus Payment for Ecosystem Services in deforestation and po-
verty in Brazil. 

## Loading the packages and data

# 1. Loading packages
```{r}
library(dplyr)
library(sf)
library(raster)
```
Note: needed to install the packages "rgdal" and "sp" to be able to use the package
"raster". And also lwgeom.

# 2. Loading the data

# 2.1 rasters

```{r}
sp_2010 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2010.tif")
sp_2011 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2011.tif")
sp_2012 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2012.tif")
sp_2013 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2013.tif")
sp_2014 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2014.tif")
sp_2015 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2015.tif")
sp_2016 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2016.tif")
sp_2017 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2017.tif")
sp_2018 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2018.tif")
sp_2019 <- raster("/Users/juliaabdalla/Desktop/Data Thesis/mapbiomas/mapbiomas-brazil-collection-50-saopaulo-2019.tif")
```

2.1.1 stacking the rasters

```{r}
rasters <- stack(sp_2010, sp_2011, sp_2012, sp_2013, sp_2014, sp_2015, sp_2016, sp_2017, sp_2018, sp_2019)
```



2.2 Shapefile

Had to change the file enconding to properly read the data. UTF8 didn't work, but
the below solution with Windows-1252 did. 
```{r}
PSA <- st_read("/Users/juliaabdalla/Desktop/Data Thesis/area_projeto_disjunta_conma/Area_projeto_disjunta_conma.shp", options = "ENCODING=WINDOWS-1252")
```

2.3 Operations: shapefile needs to have same crs as rasters

2.3.1 Retrieve rasters' crs
```{r}
crs(rasters)
```

2.3.2 Transform PSA's crs into rasters' crs
```{r}
PSA_crs <- st_transform(PSA, crs = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
```

2.3.3 Checking to see if it worked

```{r}
crs(PSA_crs)
```

This solution turned out to not give the correct output from the rasters after 
they were cropped, i.e., it looked like it cropped the wrong area. This may be
because the rasters are unprojected. So later in the code I do the reverse: match
the rasters crs to that of the PSA shapefile, which is projected. 

## Overview of the PSA dataframe

# 4. Taking a quick look at the data
```{r}
glimpse(PSA)
```
```{r}
colnames(PSA)
```

# 5. Taking a quick look at the map
```{r}
plot(PSA)
```

# 7. Inspecting one of the raster objects
```{r}

sp_2010

```

# 8. Quick look at the map
```{r}
plot(sp_2010)
```
problem: legends. This created a heatmap of the values.

# 9. Quick look at the structure 
```{r}
st_crs(sp_2010)
```
The +proj=longlat means it's unprojected. It is generally best to perform spatial analysis with layers that have a projected CRS.If it begins with +proj=longlat then the CRS is unprojected.

## Some data manipulation to be able to make the model run

# 10. Projecting the raster 

# Convert PSA to a Spatial object

```{r}
PSA_sp1 <- st_zm(PSA_crs)

PSA_sp <- as(PSA_sp1, "Spatial")
```


# Mask the rasters with the PSA area (PSA_crs) 

```{r}
rasters_masked <- mask(rasters, mask = PSA_sp)
```

# Plot rasters_masked -- this is a raster!

```{r}
plot(rasters_masked)
```


# Crop rasters with PSA_sp2. This essentially "zooms in".

```{r}
rasters_cropped <- crop(rasters, PSA_crs)
```


# Plot the cropped version and compare
plot(canopy_crop)
8. 
```{r}
plot(rasters_cropped)
```



# Extract the rasters values at the shapefile cut
landmarks_ex <- extract(canopy, landmarks_sp)

```{r}
rasters_extracted <- extract(rasters_cropped, PSA_sp2)
```


# Look at the results
landmarks_cp
landmarks_ex




