---
title: "PAs vs PES in Brazil"
author: "Julia Lunardi"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Protected Areas vs Payment for Ecosystem Services in Brazil

This is the code I am developing for my Master's thesis about the effectiveness
of Protected Areas versus Payment for Ecosystem Services in deforestation and po-
verty in Brazil. 

## Loading the packages and data

# 1. Loading packages
```{r}
library(dplyr)
library(sf)
library(raster)
```
Note: needed to install the packages "rgdal" and "sp" to be able to use the package
"raster". And also lwgeom.

# 2. Loading the data

# 2.1 rasters

```{r, echo=FALSE}
sp_2010 <- raster("file.path")
sp_2011 <- raster("file.path")
sp_2012 <- raster("file.path")
sp_2013 <- raster("file.path")
sp_2014 <- raster("file.path")
sp_2015 <- raster("file.path")
sp_2016 <- raster("file.path")
sp_2017 <- raster("file.path")
sp_2018 <- raster("file.path")
sp_2019 <- raster("file.path")
```

# 2.1.1 stacking the rasters

```{r}
rasters <- stack(sp_2010, sp_2011, sp_2012, sp_2013, sp_2014, sp_2015, sp_2016, sp_2017, sp_2018, sp_2019)
```



# 2.2 Shapefile

Had to change the file enconding to properly read the data. UTF8 didn't work, but
the below solution with Windows-1252 did. 
```{r}
PSA <- st_read("/file.path", options = "ENCODING=WINDOWS-1252")
```

# 2.3 Operations: shapefile needs to have same CRS as rasters

# 2.3.1 Retrieve rasters' crs
```{r}
crs(rasters)
```

# 2.3.2 Transform PSA's crs into rasters' crs
```{r}
PSA_crs <- st_transform(PSA, crs = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
```

# 2.3.3 Checking to see if it worked

```{r}
crs(PSA_crs)
```

## Overview of the PSA dataframe

# 3. Taking a quick look at the data

```{r}
glimpse(PSA)
```

```{r}
colnames(PSA)
```

# 4. Taking a quick look at the map

```{r}
plot(PSA)
```

# 5. Inspecting one of the raster objects

```{r}
sp_2010
```

# 6. Quick look at the map
```{r}
plot(sp_2010)
```
problem: legends. This created a heatmap of the values.

# 7. Quick look at the structure 
```{r}
st_crs(sp_2010)
```
The +proj=longlat means it's unprojected. It is generally best to perform spatial analysis with layers that have a projected CRS.If it begins with +proj=longlat then the CRS is unprojected.

## Some data manipulation to be able to make the model run


# 8. Mask the rasters with the PSA area (PSA_crs) 
-- this didnt work. Output error: Error in x$.self$finalize() : attempt to apply non-function

```{r}
rasters_masked <- mask(rasters, mask = PSA_sp)
```

# 9. Crop rasters with PSA_crs

```{r}
rasters_cropped <- crop(rasters, PSA_crs) 
```
This outputted the wrong area. 

# 10. Plot the cropped version and compare

```{r}
plot(rasters_cropped)
```
Here we see that the area cropped is not the same as the shapefile area. 


# 11. Extract the rasters values at the shapefile cut





